---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
library(eunis.habitats)
```

# eunis.habitats <a href="https://rmagno.eu/eunis.habitats/"><img src="man/figures/logo.svg" align="right" height="139" /></a>

<!-- badges: start -->
[![CRAN status](https://www.r-pkg.org/badges/version/eunis.habitats)](https://CRAN.R-project.org/package=eunis.habitats)
<!-- badges: end -->

`{eunis.habitats}` is an R data package that provides the [EUNIS Habitat
Classification](https://www.eea.europa.eu/data-and-maps/data/eunis-habitat-classification-1) in tidy format.

## Installation

Install from CRAN:

``` r
install.packages("eunis.habitats")
```

You can install the development version of `{eunis.habitats}` like so:

``` r
# install.packages("remotes")
remotes::install_github("ramiromagno/eunis.habitats")
```

## EUNIS classifications

There are four EUNIS classifications included in the one single data set `eunis_habitats`, namely:

1. EUNIS classification from 2007, revised in 2012.
2. EUNIS marine classification from 2019
3. EUNIS marine classification from 2022
4. EUNIS terrestrial classification from 2021

For example, to access the marine habitats classified according to EUNIS revision of 2022:

```{r}
subset(eunis_habitats, classification == "EUNIS_M_2022")
```

Check the documentation for more details: `?eunis_habitats`.

## Crosswalks (mapping of habitat codes)

From EUNIS 2012 to EUNIS Marine 2022:

```{r}
crosswalk(
  code = c("A3.4", "A3.5"),
  from = "EUNIS_2012",
  to = "EUNIS_M_2022",
  unnest = TRUE
)
```

From EUNIS Marine 2019 to EUNIS 2012:

```{r}
crosswalk(
code = c("MH152", "MH2331"),
from = "EUNIS_M_2019",
to = "EUNIS_2012",
unnest = TRUE
)
```

From EUNIS Marine 2022 to the European Red List of Habitats:

```{r}
crosswalk(
code = c("MH152", "MH2331", "MA146", "MD55"),
from = "EUNIS_M_2022",
to = "RL",
unnest = TRUE
)
```

From EUNIS Marine 2019 to Habitats Directive Annex I:

```{r}
crosswalk(
code = c("M", "MA1", "MA11", "MA12"),
from = "EUNIS_M_2019",
to = "Annex_I", unnest = TRUE
)
```

## Original data source

The original data is by [EUNIS Habitat
Classification](https://www.eea.europa.eu/data-and-maps/data/eunis-habitat-classification-1).
