---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
library(eunis.habitats)
```

# eunis.habitats

<!-- badges: start -->
[![CRAN status](https://www.r-pkg.org/badges/version/eunis.habitats)](https://CRAN.R-project.org/package=eunis.habitats)
<!-- badges: end -->

`{eunis.habitats}` is an R data package that provides the [EUNIS Habitat
Classification](https://www.eea.europa.eu/data-and-maps/data/eunis-habitat-classification-1) in tidy format.

## Installation

You can install the development version of `{eunis.habitats}` like so:

``` r
# install.packages("remotes")
remotes::install_github("ramiromagno/eunis.habitats")
```

## EUNIS classifications

There are four EUNIS classifications included in the one data set
`eunis_habitats` are:

1. EUNIS classification from 2007, revised in 2012.
2. EUNIS marine classification from 2019
3. EUNIS marine classification from 2022
4. EUNIS terrestrial classification from 2021

Check the documentation for more details: `?eunis_habitats`.

## EUNIS 2012

The full list of EUNIS habitats 2012: codes, scientific names and revised descriptions. The habitat descriptions of the EUNIS classification 2007 were revised in 2012. The 2007 habitat types were not changed in the 2012 description revision which mostly replaced Palaearctic or UK Marine habitat classification codes used in habitat descriptions at levels 5 and below with their EUNIS classification equivalents. In 2019 the classification was further amended to include two new habitats of the revised Resolution 4 of Bern Convention as adopted at the 38th Standing Committee meeting, November 2018. The two habitats are G3.4G Pinus sylvestris forest on chalk in the steppe zone and X36 Depressions (pody) of the Steppe zone.

```{r EUNIS_2012}
subset(eunis_habitats, classification == "EUNIS_2012")
```

## EUNIS marine 2019

A revision of EUNIS marine habitat classification was published in 2019. From June 2021 the classification was complemented with crosswalks to Habitats Directive Annex I and to European Red List of Habitats. An updated version was published in March 2022. Users should change to the 2022 version at their earliest convenience. 

```{r EUNIS_M_2019}
subset(eunis_habitats, classification == "EUNIS_M_2019")
```

## EUNIS marine 2022

The review of the marine component of the EUNIS habitat classification was initiated in 2014. Marine benthic habitats, marine pelagic and marine ice associated habitats are separated into three distinct groups, each with a separate classification structure. The first major division in the benthic marine part of the EUNIS classification is based on major biological zones (related to depth) and substrate type. Level 3 of the classification reflects the main biogeographical regions of Europe’s seas based on their distinct combinations of salinity and temperature regimes (Arctic, Baltic, Atlantic, Mediterranean and Black Sea). A first review was published in 2019 and an update to this version concerning mostly the Atlantic regional sea is available since March 2022. Crosswalks to Habitats Directive Annex I and to European Red List of Habitats are available while crosswalks to EUNIS marine habitats of version 2012 for the regional seas apart from the Atlantic need to be revisited.

```{r EUNIS_M_2022}
subset(eunis_habitats, classification == "EUNIS_M_2022")
```

## EUNIS terrestrial 2021

The review of the terrestrial component of the EUNIS habitat classification was initiated in 2015. The review concerns the groups of coastal habitats, grasslands, heathland, forest, sparsely vegetated and vegetated man-made habitats. The review was updated in January 2023 with some corrections to the existing groups and with the inclusion of wetlands and additional crosslinks. The remaining groups will be revised and published at a later stage. The classification includes cross-walks at level 3 to Habitats Directive Annex I, to European Red List of Habitats, to Bern Convention Resolution 4 habitats, to MAES and IUCN ecosystems, to Corine Land Cover classes and to the Euroveg Checklist 2016 Syntaxa. The groups of forest and heathland include also crosswalks to an earlier revision of these groups published in 2017. Habitats at level 3 are complemented with lists of characteristic species identified from the EVA database. 

```{r EUNIS_T_2021}
subset(eunis_habitats, classification == "EUNIS_T_2021")
```

## Original data source

The original data is by [EUNIS Habitat
Classification](https://www.eea.europa.eu/data-and-maps/data/eunis-habitat-classification-1).
